Combinatorics of the geometry of Wilson loop diagrams II: Grassmann necklaces, dimensions, and denominators

### 2019-09-03 GWLDCII.tex
read to end. added too .bib fil

Generally speaking the propagators in a Wilson loop diagram are undirected, so $(i,j)$ and $(j,i)$ represent the same propagator. If we need to impose a direction on a propagator $p = (i,j)$, for example in order to consider the region ``inside'' or ``outside'' of $p$, we write $(p,<_i)$ to indicate that $p$ is directed from edge $i$ to edge $j$, and $(p,<_j)$ for the opposite direction. {\color{violet} Note to consider while reading: Might it be notationally easier and/or clearer to talk about ordering the propagators as clockwise or counter clockwise? If so, are the benefits of this new notation worth the rewriting?}

@@

\section{Poles of Wilson Loop Integrals}\label{sec poles} \sanote{start here}

In this section, we investigate the poles of the integrand for $I(W)$, that is $R(W)$ as defined in Definition~\ref{def R(W)}. It is conjectured that for a fixed number of propagators and a given set of external twistors $\cZ$, [cite]\note{A cite still waiting} \bas \sum_{W | |\cP| = k} I(W)(\cZ) <\infty \;.\eas In otherwords, all poles of the intergrals must cancel. As a step towards this, in this section, we prove some results about $R(W)$.

The results of Section~\ref{sec GN algorithm} allow us to relate the position of propagators in a Wilson loop diagram $W$ to minors of $C(W)$, $R(W)$.

The main result of this section is Theorem~\ref{thm denom}, which expresses the denominator $R(W)$ in terms of the Grassmann necklace of $W$. This simplifies the computation of $R(W)$ and allows us to directly relate the poles of the integral to the combinatorics of the diagram.

We first give an algorithm which extracts the required minors from the Grassmann necklace.

\begin{algorithm}\label{alg WLD to denom via GN}
Let $W = (\cP,n)$ be a Wilson loop diagram, and let $C(W)$ be the matrix of $W$ as defined in \eqref{C(W) dfn} with a fixed ordering on $\cP$. (see Section~\ref{section WLD background}).
\begin{itemize}
  \item For each $i \in [n]$, we construct a factor $r_i$ as follows:
    \begin{itemize}
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i-1$st Grassmann necklace element, $I_{i-1}$.  (By convention set $I_{0}=I_n$.)  Write $I_{i-1}(p)$ for the vertex associated to propagator $p$ under this bijection.
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i$th Grassmann necklace element, $I_{i}$.  Write $I_{i}(p)$ for the vertex associated to propagator $p$ under this bijection.
      \item Let $S_i = \{p \in \cP \ | \ I_{i-1}(p) \neq I_i(p)\}$. (By convention, set $I_{0} = I_n$.)
	  %\item Write $\Delta_{I_i}$ for the determinant of the $k \times k$ minor of $C(W)$ with columns indexed by $I_i$.
      \item Let $r_i$ be the determinant of the $|S_i|\times |S_i|$ minor of $C(W)$ with rows indexed by $S_i$ and columns indexed by $I_i(S_i)$.  %$\Delta_{I_i}$ with all variables from rows associated to $p\not\in S_i$ set to $1$.
    \end{itemize}
  \item Define $R = \prod_{i=1}^n r_i$.
\end{itemize}
\end{algorithm}

@@

As we discuss the algorithm it will also be useful to have the notation $\Delta_{I_i}$ for the determinant of the $k\times k$ minor of $C(W)$ with columns indexed by $I_i$.

As we take each $r_i$ to simply be the determinant of a particular submatrix of $C(W)$, the sign of each $r_i$ is well-defined.  Before proving the main results of this section, we given an illustrative example. Below, we show that $R$ is equal to the denominator of the Wilson loop diagram as defined in Definition~\ref{def R(W)}, and is the radical of the product $\prod_{i=1}^n \Delta_{I_i}$.
%  However, we will not be interested in these signs as they only contribute an overall sign to $R$ and our main theorem in this section (Theorem~\ref{thm denom}) will simply be to show that $R$ is equal up to a $\mathbb{Q}$-multiple to the denominator of the Wilson loop diagram (see Definition~\ref{def R(W)}).

@@

\begin{eg}
Consider the Wilson loop diagram in Figure~\ref{fig R eg}. Assigning propagators $p$, $q$, $r$ to rows $1,2,3$ respectively, we obtain the matrix
\[
C(W) = \begin{bmatrix} a & b & 0 & 0 & 0 & c & d \\ e & f & 0 & 0 & g & h & 0 \\ i & j & 0 & k & l & 0 & 0 \end{bmatrix}
\]
The Grassmann necklace of this diagram is 
\begin{gather*}I_1 = \{1,2,4\}, I_2 = \{2,4,5\}, I_3 = \{4,5,6\}, I_4=\{4,5,6\},\\ I_5=\{5,6,7\}, I_6 = \{6,7,1\}, I_7=\{7,1,2\}. \end{gather*}  
Figure~\ref{fig R eg} indicates the pairings between propagators and vertices for each $i \in [1,7]$.

From $I_1$ to $I_2$, the propagators $p$ and $q$ change which vertices they contribute. That is $I_1(p) = 1$, $I_1(q)= 2$ while $I_2(p) = 2$, $I_2(q)= 5$. However, $r$ contributes  4 to both $I_1$ and $I_2$, so $S_2 = \{p,q\}$.  Then
\[
\Delta_{I_2}=\det\begin{bmatrix} b & 0 & 0 \\ f & 0 & g \\ j & k & l \end{bmatrix} = kgb, \qquad r_2 = \det\begin{bmatrix} b & 0 \\ f & g \end{bmatrix} = gb.
\] 
Continuing likewise, we get $r_3 = c$, $r_4=1$ (since $I_4 = I_3$), $r_5 = lhd$, and $r_6 = i$.

At $I_7$ the situation we see for the first time an irreducible quadratic factor: we have $S_7 = \{q,s\}$ with $I_7(S_7) = \{ 1,2\}$, $\Delta_{I_7} = d(ej-fi)$ and $r_7 = ej-fi$.

@@

\begin{prop}\label{prop alg gives rad}
  With notation as in Algorithm~\ref{alg WLD to denom via GN} we have the following:
  \begin{enumerate}
    \item Each $\Delta_{I_i}$ is homogeneous, as is each $r_i$.
    \item Each $\Delta_{I_i}$ splits into linear and quadratic factors.  All linear factors of  $\Delta_{I_i}$ are single variables and all irreducible quadratic factors are $2\times 2$ determinants of single variables.
    \item Quadratic factors in $\Delta_{I_i}$ arise precisely when propagators $p$ and $q$ are supported on a common edge $(a,b)$ with $I_i(p)=a$ and $I_i(q)=b$.
    \item $r_i$ divides $\Delta_{I_i}$.
    \item The ideal generated by $R$ is the radical of the ideal generated by $\prod_{i=1}^{n}\Delta_{I_i}$.
  \end{enumerate}
\end{prop}

\begin{proof}
\begin{enumerate}
    \item The nonzero entries of $C(W)$ are independent indeterminates and so every $i\times i$ minor of $C(W)$ is either homogeneous of degree $i$ or is $0$.  Thus each $\Delta_{I_i}$ and each $r_i$ is homogeneous.  %Furthermore, each row contributes one factor to each term in the expansion of $\Delta_{I_i}$ so the result of setting the variables from a subset of rows to $1$ is still homogeneous.  Thus each $r_i$ is homogeneous.
    \item Using the expression for the determinant as a sum over permutations along with Corollary~\ref{lem basis as perm} we see that $\Delta_{I_i}$ is a sum over bijections between $I_i$ and $\mathcal{P}$.  The nonzero terms in this sum are precisely those bijections such that each propagator is associated to one of its supporting vertices in $I_i$, since only those locations in $C(W)$ are nonzero.  Since the nonzero entries of $C(W)$ are independent there can be no cancellation between terms in this expansion.

@@

Expanding out $fg$ yields a signed sum of monomials. Since there is no cancellation between terms, this means that the full expansion over permutations of $\Delta_{I_i}$ contains no other nonzero terms and hence no other variables.  Therefore $\Delta_{I_i}$ is equal to the determinant of the matrix obtained by taking the submatrix of $C(W)$ with columns indexed by $I_i$ and setting any variables not appearing in $\Delta_{I_i}$ to $0$.  This new matrix is, up to permutations of rows and columns, a block matrix with one block for $\mathcal{P}'$ and $J$ and the other block for the complements.  Thus its determinant, and hence also $\Delta_{I_i}$, is the product of the minors for these two blocks.  By considering which variables appear, these two factors must also be $f$ and $g$, and so in particular $f$ is the minor of $C(W)$ associated to $\mathcal{P}'$ and $J$.  This proves the claim.

@@

All that remains is to prove that $\Delta_{I_i}$ has no irreducible factors of degree 3 or more.  Suppose for a contradiction that $f$ is a factor of $\Delta_{I_i}$ of degree $\geq 3$. Without loss of generality\footnote{This is without loss of generality as by removing the propagators which come before those contributing to $f$ in the order given by the construction of $I_i$ and changing $i$ to be the first vertex which contributes to $f$, we obtain an admissible diagram for which $f$ still divides $\Delta_{I_i}$ but also $i\in I_i$ and
$i$ contributes to $f$, and this diagram can be used in place of $W$.} we may assume $i\in I_i$ and $i$ contributes to $f$.
Finally, we can suppose that $W$ is minimal in number of propagators with an irreducible factor of degree $\geq 3$.

Let $p$ be the propagator such that $I_i(p) = i$. There are two cases to consider, depending on which edge $p$ is supported on.  These are illustrated in Figure~\ref{fig no big factors}

\begin{figure}
\[
\begin{tikzpicture}[baseline=(current bounding box.east),rotate=-20]
\begin{scope}
	\drawWLDfragment[8]{3}{0.4} 
	%\drawnumberspartial % use this to debug
	\newnode[left]{1}{i-1}
	\newnode[below]{2}{i}
	\newnode[below]{7}{m}
	\begin{scope}
		\clipcenterarc(0,0)(\startpoint-10:\endpoint+10:\radius)
		\newpropbend{1.7}{7.2}{60}
		\draw[smallpropagator,blue] (\zero+7.8*\step:\radius*1.1) to[bend left = 55] (\zero + 10*\step:\radius*1.1);
		\node at (\zero + 1.2*\step:\radius*0.5) {\footnotesize $p$};
	\end{scope}
	\node[align = center,black] at (\zero + \step*4.5:\radius*0.65) {\em \footnotesize propagators \\ \footnotesize \em inside $p$};
	\draw[->,shorten >=2pt,thick,gray] (\zero + 1.6*\step:\radius) to[bend left = 45] (\zero + 1*\step:\radius); %p
	\draw[->,shorten >=2pt,thick,gray] (\zero + 7.9*\step:\radius) to[bend left = 45] (\zero + 7*\step:\radius); 
	\node[align = center,black] at (\zero + 6.8*\step:1.5*\radius) {\em \footnotesize If $m \in I_i$ then \\ \em \footnotesize it was contributed by \\ \em \footnotesize the blue propagator};
\end{scope}
\end{tikzpicture}
\qquad 
\begin{tikzpicture}[baseline=(current bounding box.east),rotate=-20]
	\begin{scope}
	\drawWLDfragment[8]{3}{0.4}  
	%\drawnumberspartial % use this to debug
	\newnode[left]{1}{i}
	\newnode[below]{2}{i+1}
	\newnode[below]{7}{j}
	\centerarc[red,line width = 2pt,line cap = round](0,0)(\zero + 1.5*\step:\zero+1.56*\step:\radius);
	\draw[->,shorten >=2pt,thick,gray] (\zero + 1.45*\step:\radius) to[bend left = 45] (\zero + 1*\step:\radius); %p
	\draw[->,shorten >=2pt,thick,gray] (\zero + 1.61*\step:\radius) to[bend right = 45] (\zero + 2*\step:\radius); %q 
	\draw[->,shorten >=2pt,thick,gray] (\zero + 7.7*\step:\radius) to[bend left = 45] (\zero + 7*\step:\radius); %blue
	\begin{scope}
		\clipcenterarc(0,0)(\startpoint-10:\endpoint+10:\radius)
		\newpropbend{1.6}{7.2}{40} % q
		\draw[smallpropagator] (\zero+1.5*\step:\radius*1.1) to[bend right = 40] (\zero  - 3*\step:\radius*1.1); % p
		\draw[smallpropagator,blue] (\zero+7.6*\step:\radius*1.1) to[bend left = 55] (\zero + 10*\step:\radius*1.1);
		\node at (\zero + 5*\step:\radius*0.3) {\footnotesize $q$};
		\node at (\zero + 1*\step:\radius*0.6) {\footnotesize $p$};
	\end{scope}
	\node[align = center,black] at (\zero + \step*4.5:\radius*0.75) {\em \footnotesize propagators \\ \footnotesize \em inside $q$};
	\node[align = center] at (\zero + 1.5*\step:\radius*1.5) {\footnotesize \em No \\ \footnotesize \em propagators \\ \footnotesize \em between \\ \footnotesize \em $p$ and $q$ \\ \footnotesize \em here};
	\node[align = center,black] at (\zero + 6.8*\step:1.5*\radius) {\em \footnotesize If $j \in I_i$ then \\ \em \footnotesize it was contributed by \\ \em \footnotesize the blue propagator};
\end{scope}
\end{tikzpicture}
\]
  \caption{The two cases in the proof that no factors of $\Delta_{I_i}$ have degree 3 or more.}\label{fig no big factors}
\end{figure}

\textbf{Case 1}: Suppose $p$ has one end on the edge $(i-1, i)$.  Thus $p$ is supported on $(i-1, i, m, m+1)$ for some $m >_i i$, and $I_{i+1}(p) = m$ by Lemma~\ref{vertex cyclic int lem}.

Let $S = \cP_{in}(p)$ be the set of propagators inside $p$ in the $<_i$ order. $I_i$ and $I_{i+1}$ can only differ once $p$ contributes to $I_{i+1}$, so $I_i(q) = I_{i+1}(q)$ for each $q \in S \backslash \{p\}$. Thus if a propagator contributes $m$ in $I_i$ then it must lie outside $p$.

If neither $m$ nor $m+1$ appear in $I_i$ then by Corollary~\ref{no coloops} $V(p) \cap I_i = \{i\}$, and so the row of $p$ in the matrix of $\Delta_{I_i}$ has only one nonzero entry; hence $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$, which is a contradiction to the fact that $i$ contributes to $f$.  So we must have at least one of $m$ and $m+1$ in $I_i$.  However, all propagators in $S$ are mapped by the function $I_{i}(\cdot)$ to vertices strictly before $m$, so the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the $|S|\times |S|$ matrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. No other propagators can be supported on these vertices since all other propagators are outside of $p$, and $p$ is the first propagator supported at $i$; this explains the zero block.  Therefore $\Delta_{I_i} = \det A \det C$, and both factors are nontrivial since at least one of $m$ and $m+1$ appear in $I_i$.  Since $p\in \cP_{in}(p)$ we have $f|\det A$ and if we remove the propagator outside of $p$ that contributes $m$ or $m+1$, we get a diagram with fewer propagators for which $\det A | \Delta_{I_i}$, hence $f|\Delta_{I_i}$.  This contradicts the minimality of our choices.

\textbf{Case 2}: Suppose $p$ has one end on the edge $(i, i+1)$.  If no other propagators are supported on $i$ then the column of $C(W)$ corresponding to vertex $i$ has only one nonzero entry in it, and so $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$; as above, this is a contradiction.  Since $I_i(p) = i$ there can be no propagator supported on $i$ with an end on the edge $(i-1, i)$.  Thus to have another propagator supported on $i$ there is at least one other propagator on edge $(i,i+1)$.  Further since since $I_i(p)=i$, one of these other propagators contributes $i+1$ in $I_i$, so we can take $q$ to be the propagator such that $I_i(q)=i+1$.  We know that $q$ has one end on the edge $(i, i+1)$ and is adjacent to $p$ on that edge in the counterclockwise direction (see Figure \ref{fig no big factors}).  Write $(i, i+1, j, j+1)$ for the support of $q$.  The situation for $q$ is very similar to case 1: in particular, we have $I_{i+1}(q) = j$ by Lemma~\ref{vertex cyclic int lem} and so if $j\in I_i$ then the propagator which contributes $j$ is outside of $q$.

Similarly to Case 1, let $S = \cP_{in}(q) \cup \{p\}$ in the $<_i$ order. Then all propagators in $S$ are mapped by $I_i(\cdot)$ to vertices strictly before $j$ and no other propagators are supported on vertices strictly before $j$.  Thus the matrix giving $\Delta_{I_i}$ has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the submatrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. Again two things can now happen.  If some vertex $j$ or larger (with respect to $>_i$) belongs to $I_i$ then $B$ and $C$ are at least one column wide, and so the block form of the matrix gives a nontrivial factorization of $\Delta_{I_i}=\det A\det B$.  This yields a contradiction as in Case 1: $f|\det A$ and removing a propagator not in $S$ gives a smaller diagram with a factor of degree $\geq 3$ which contradicts our minimality assumption.\note{I rephrased this last sentence quite a bit because earlier things had gotten rephrased}

>>>>

Generally speaking the propagators in a Wilson loop diagram are undirected, so $(i,j)$ and $(j,i)$ represent the same propagator. If we need to impose a direction on a propagator $p = (i,j)$, for example in order to consider the region ``inside'' or ``outside'' of $p$, we write $(p,<_i)$ to indicate that $p$ is directed from edge $i$ to edge $j$, and $(p,<_j)$ for the opposite direction.

@@

\section{Poles of Wilson Loop Integrals}\label{sec poles}

In this section, we investigate the poles of the integrand for $I(W)$, that is $R(W)$ as defined in Definition~\ref{def R(W)}. Since SYM $N=4$ is a finite theory \cite{Alday:2007hr, ParkeTaylorids}, the scattering amplitudes (and thus the holomorphic Wilson loops $\,\mathfrak{W}_{k,n}$) are finite. That is, for a fixed set of external particles $\cZ$, \bas \sum_{W \text{ s.t. } |\cP| = k} I(W)(\cZ) <\infty \;.\eas In otherwords, all poles of the intergrands must cancel. As a step towards this, in this section, we prove some results about the structure of the denominator of the integrand, $R(W)$.

The results of Section~\ref{sec GN algorithm} allow us to relate the \hlfix{position of propagators}{ending edges?} in a Wilson loop diagram $W$ to minors of $C(W)$, $R(W)$. \sanote{Do we need to fix an ordering of rows of $C(W)$, or does that come out in the wash?}

The main result of this section is Theorem~\ref{thm denom}, which expresses the denominator $R(W)$ in terms of the Grassmann necklace of $W$. \sanote{Put something about this in the motivation for why WLD $\rightarrow$ GN algorithm is useful}This simplifies the computation of $R(W)$ and allows us to directly relate the poles of the integral to the combinatorics of the diagram.

We first give an algorithm which extracts the required minors of $C(W)$ from the Grassmann necklace.

\begin{algorithm}\label{alg WLD to denom via GN}
Let $W = (\cP,[n])$ be a Wilson loop diagram, and let $C(W)$ be the matrix of $W$ as defined in \eqref{C(W) dfn} with a fixed ordering on $\cP$. (see Section~\ref{section WLD background}).
\begin{itemize}
  \item For each $i \in [n]$, we construct a factor $r_i$ as follows:
    \begin{itemize}
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i-1$st Grassmann necklace element, $I_{i-1}$.  (By convention set $I_{0}=I_n$.)  Write $I_{i-1}(p)$ for the vertex associated to propagator $p$ under this bijection.
      %\item Use Algorithm~\ref{alg:put GN on WLD} to obtain a bijection between the propagators and the $i$th Grassmann necklace element, $I_{i}$.  Write $I_{i}(p)$ for the vertex associated to propagator $p$ under this bijection.
      \item Let $S_i = \{p \in \cP \ | \ I_{i-1}(p) \neq I_i(p)\}$. (By convention, set $I_{0} = I_n$.)
	  %\item Write $\Delta_{I_i}$ for the determinant of the $k \times k$ minor of $C(W)$ with columns indexed by $I_i$.
      \item Let $r_i$ be the determinant of the $|S_i|\times |S_i|$ minor of $C(W)$ with rows indexed by $S_i$ and columns indexed by $I_i(S_i)$.  {\color{violet} If $S_i = \emptyset$, set $r_i = 1$}\sanote{a version of this was commented out, put back in}%$\Delta_{I_i}$ with all variables from rows associated to $p\not\in S_i$ set to $1$.
    \end{itemize}
  \item Define $R = \prod_{i=1}^n r_i$.
\end{itemize}
\end{algorithm}

@@

As we discuss the algorithm it will also be useful to have the notation $\Delta_{I_i}$ for the determinant of the $k\times k$ minor of $C(W)$ with columns indexed by $I_i$. {\color{violet} In particular, one summand of this determinant is given by $\prod_{p \in \cP}x_{p, I_i(p)}$. The other summands are (up to sign) given by other bijections between the propagator set $\cP$ and the vertex set $I_i$.}

As we take each $r_i$ to simply be the determinant of a particular submatrix of $C(W)$, the sign of each $r_i$ is well-defined {\color{violet} (after an ordering on the propagators have been fixed)}.  Before proving the main results of this section, we given an illustrative example. Below, we show that $R$ is equal to the denominator of the Wilson loop diagram as defined in Definition~\ref{def R(W)}, and is the radical of the product $\prod_{i=1}^n \Delta_{I_i}$.
%  However, we will not be interested in these signs as they only contribute an overall sign to $R$ and our main theorem in this section (Theorem~\ref{thm denom}) will simply be to show that $R$ is equal up to a $\mathbb{Q}$-multiple to the denominator of the Wilson loop diagram (see Definition~\ref{def R(W)}).

@@

\begin{eg}
Consider the Wilson loop diagram in Figure~\ref{fig R eg}. Assigning propagators $p$, $q$, $r$ to rows $1,2,3$ respectively, we obtain the matrix
\[
C(W) = \begin{bmatrix} a & b & 0 & 0 & 0 & c & d \\ e & f & 0 & 0 & g & h & 0 \\ i & j & 0 & k & l & 0 & 0 \end{bmatrix}
\]
The Grassmann necklace of this diagram is 
\begin{gather*}I_1 = \{1,2,4\}, I_2 = \{2,4,5\}, I_3 = \{4,5,6\}, I_4=\{4,5,6\},\\ I_5=\{5,6,7\}, I_6 = \{6,7,1\}, I_7=\{7,1,2\}. \end{gather*}  
Figure~\ref{fig R eg} indicates \st{the pairings between propagators and vertices for each }{\color{violet} indicates which vertex is contributed to $I_i$ by each propagator for each} $i \in [1,7]$.

From $I_1$ to $I_2$, the propagators $p$ and $q$ change which vertices they contribute. That is $I_1(p) = 1$, $I_1(q)= 2$ while $I_2(p) = 2$, $I_2(q)= 5$. However, $r$ contributes  4 to both $I_1$ and $I_2$, so $S_2 = \{p,q\}$.  Then
\[
\Delta_{I_2}=\det\begin{bmatrix} b & 0 & 0 \\ f & 0 & g \\ j & k & l \end{bmatrix} = kgb, \qquad r_2 = \det\begin{bmatrix} b & 0 \\ f & g \end{bmatrix} = gb.
\] 
Continuing likewise, we get $r_3 = c$, $r_4=1$ \sanote{see bit added to Alg 5.1 above}(since $I_4 = I_3$), $r_5 = lhd$, and $r_6 = i$.

At $I_7$ we see for the first time an irreducible quadratic factor: we have $S_7 = \{q,s\}$ with $I_7(S_7) = \{ 1,2\}$, $\Delta_{I_7} = d(ej-fi)$ and $r_7 = ej-fi$.

@@

\begin{prop}\label{prop alg gives rad}
  With notation as in Algorithm~\ref{alg WLD to denom via GN} we have the following:
  \begin{enumerate}
    \item Each $\Delta_{I_i}$ is homogeneous, as is each $r_i$.
    \item Each $\Delta_{I_i}$ splits into linear and quadratic factors.  All linear factors of  $\Delta_{I_i}$ are single variables and all irreducible quadratic factors are $2\times 2$ determinants of single variables.
    \item Quadratic factors in $\Delta_{I_i}$ arise precisely when propagators $p$ and $q$ are supported on a common edge \st{$(a,b)$}{\color{violet}$(a,a+1)$} with $I_i(p)=a$ and \st{$I_i(q)=b$}{\color{violet}$I_i(q)=a+1$}.\sanote{would like to make this change throughout}
    \item $r_i$ divides $\Delta_{I_i}$.
    \item The ideal generated by $R$ is the radical of the ideal generated by $\prod_{i=1}^{n}\Delta_{I_i}$.
  \end{enumerate}
\end{prop}

\begin{proof}
\begin{enumerate}
    \item The nonzero entries of $C(W)$ are independent indeterminates and so every $i\times i$ minor of $C(W)$ is either homogeneous of degree $i$ or is $0$.  Thus each $\Delta_{I_i}$ and each $r_i$ is homogeneous.  %Furthermore, each row contributes one factor to each term in the expansion of $\Delta_{I_i}$ so the result of setting the variables from a subset of rows to $1$ is still homogeneous.  Thus each $r_i$ is homogeneous.
    \item Using the expression for the determinant as a sum over permutations along with Corollary~\ref{lem basis as perm} we see that $\Delta_{I_i}$ is a sum over bijections between $I_i$ and $\mathcal{P}$.  The nonzero terms in this sum are precisely those bijections such that each propagator is associated to one of its supporting vertices in $I_i$, namely $I_i(p)$, since only those locations in $C(W)$ are nonzero.  Since the nonzero entries of $C(W)$ are independent there can be no cancellation between terms in this expansion.

@@

Expanding out $fg$ yields a signed sum of monomials. Since there is no cancellation between terms, this means that the full expansion over permutations of $\Delta_{I_i}$ contains no other nonzero terms and hence no other variables.  Therefore $\Delta_{I_i}$ is equal to the determinant of the matrix obtained by taking the submatrix of $C(W)$ with columns indexed by $I_i$ and setting any variables not appearing in $\Delta_{I_i}$ to $0$.  This new matrix is, up to permutations of rows and columns, a block matrix with one block for $\mathcal{P}'$ and $J$ and the other block for the complements.  Thus its determinant, and hence also $\Delta_{I_i}$, is the product of the minors for these two blocks.  By considering which variables appear, these two factors must also be $f$ and $g$, and so in particular $f$ is the minor of $C(W)$ associated to $\mathcal{P}'$ and $J$.  This proves the claim. \sanote{I know there was some contention over this paragraph. Having read it carefully, I want to iterate that I like this now.}

@@

All that remains is to prove that $\Delta_{I_i}$ has no irreducible factors of degree 3 or more.  Suppose for a contradiction that $f$ is a factor of $\Delta_{I_i}$ of degree $\geq 3$. Without loss of generality\footnote{This is without loss of generality as by removing the propagators which come before those contributing to $f$ in the order given by $I_i$ and changing $i$ to be the first vertex which contributes to $f$, we obtain an admissible diagram for which $f$ still divides $\Delta_{I_i}$ but also $i\in I_i$ and
$i$ contributes to $f$, and this diagram can be used in place of $W$.} we may assume $i\in I_i$ and $i$ contributes to $f$.
Finally, we can suppose that $W$ is minimal in number of propagators with an irreducible factor of degree $\geq 3$. \sanote{can we break footnote inot a couple sentences?}

Let $p$ be the propagator such that $I_i(p) = i$. There are two cases to consider, when $p$ is supported on the edge $i-1$ or the edge $i$.  These are illustrated in Figure~\ref{fig no big factors}

\begin{figure}
\[
\begin{tikzpicture}[baseline=(current bounding box.east),rotate=-20]
\begin{scope}
	\drawWLDfragment[8]{3}{0.4} 
	%\drawnumberspartial % use this to debug
	\newnode[left]{1}{i-1}
	\newnode[below]{2}{i}
	\newnode[below]{7}{m}
	\begin{scope}
		\clipcenterarc(0,0)(\startpoint-10:\endpoint+10:\radius)
		\newpropbend{1.7}{7.2}{60}
		\draw[smallpropagator,blue] (\zero+7.8*\step:\radius*1.1) to[bend left = 55] (\zero + 10*\step:\radius*1.1);
		\node at (\zero + 1.2*\step:\radius*0.5) {\footnotesize $p$};
	\end{scope}
	\node[align = center,black] at (\zero + \step*4.5:\radius*0.65) {\em \footnotesize propagators \\ \footnotesize \em inside $p$};
	\draw[->,shorten >=2pt,thick,gray] (\zero + 1.6*\step:\radius) to[bend left = 45] (\zero + 1*\step:\radius); %p
	\draw[->,shorten >=2pt,thick,gray] (\zero + 7.9*\step:\radius) to[bend left = 45] (\zero + 7*\step:\radius); 
	\node[align = center,black] at (\zero + 6.8*\step:1.5*\radius) {\em \footnotesize If $m \in I_i$ then \\ \em \footnotesize it was contributed by \\ \em \footnotesize the blue propagator};
\end{scope}
\end{tikzpicture}
\qquad 
\begin{tikzpicture}[baseline=(current bounding box.east),rotate=-20]
	\begin{scope}
	\drawWLDfragment[8]{3}{0.4}  
	%\drawnumberspartial % use this to debug
	\newnode[left]{1}{i}
	\newnode[below]{2}{i+1}
	\newnode[below]{7}{j}
	\centerarc[red,line width = 2pt,line cap = round](0,0)(\zero + 1.5*\step:\zero+1.56*\step:\radius);
	\draw[->,shorten >=2pt,thick,gray] (\zero + 1.45*\step:\radius) to[bend left = 45] (\zero + 1*\step:\radius); %p
	\draw[->,shorten >=2pt,thick,gray] (\zero + 1.61*\step:\radius) to[bend right = 45] (\zero + 2*\step:\radius); %q 
	\draw[->,shorten >=2pt,thick,gray] (\zero + 7.7*\step:\radius) to[bend left = 45] (\zero + 7*\step:\radius); %blue
	\begin{scope}
		\clipcenterarc(0,0)(\startpoint-10:\endpoint+10:\radius)
		\newpropbend{1.6}{7.2}{40} % q
		\draw[smallpropagator] (\zero+1.5*\step:\radius*1.1) to[bend right = 40] (\zero  - 3*\step:\radius*1.1); % p
		\draw[smallpropagator,blue] (\zero+7.6*\step:\radius*1.1) to[bend left = 55] (\zero + 10*\step:\radius*1.1);
		\node at (\zero + 5*\step:\radius*0.3) {\footnotesize $q$};
		\node at (\zero + 1*\step:\radius*0.6) {\footnotesize $p$};
	\end{scope}
	\node[align = center,black] at (\zero + \step*4.5:\radius*0.75) {\em \footnotesize propagators \\ \footnotesize \em inside $q$};
	\node[align = center] at (\zero + 1.5*\step:\radius*1.5) {\footnotesize \em No \\ \footnotesize \em propagators \\ \footnotesize \em between \\ \footnotesize \em $p$ and $q$ \\ \footnotesize \em here};
	\node[align = center,black] at (\zero + 6.8*\step:1.5*\radius) {\em \footnotesize If $j \in I_i$ then \\ \em \footnotesize it was contributed by \\ \em \footnotesize the blue propagator};
\end{scope}
\end{tikzpicture}
\]
  \caption{The two cases in the proof that no factors of $\Delta_{I_i}$ have degree 3 or more.{\color{violet}First picture: I thought $I_i(p) = i$. Doesn't the arrow say that $I_i(p) = i-1$?}}\label{fig no big factors}
\end{figure}

\textbf{Case 1}: Suppose $p = (i-1, m)$. Then $p$ has one end on $i-1^{th}$ edge and $I_{i+1}(p) = m$ by Lemma~\ref{vertex cyclic int lem}.  \sanote{again, it would be nice if this could be read off from a statement of a result somehwere}.

Let $S = \cP_{in}(p)$ be the set of propagators inside $p$ in the $<_i$ order. {\color{violet} By Remark \ref{rmk algorithm locally same},} $I_i$ and $I_{i+1}$ can only differ once $p$ contributes to $I_{i+1}$, so $I_i(q) = I_{i+1}(q)$ for each $q \in S \backslash \{p\}$. Thus if a propagator contributes $m$ in $I_i$ then it must lie outside $p$.

\hlfix{By Remark {\ref{clockwise ordering rem}}, if $I_i(q) = m$, then it must be the clockwise most propagator supported on $m$ that has not yet contributed to $I_i$, i.e. is lies in $\cP_{out}(p, <_i)$.}{more direct than above para. Alt. only. Don't need to adopt. (keep dfn of $S$.)}

If neither $m$ nor $m+1$ appear in $I_i$ then by Corollary~\ref{no coloops} $V(p) \cap I_i = \{i\}$, and so the row of $p$ in the matrix of $\Delta_{I_i}$ has only one nonzero entry; hence $\Delta_{I_i}$ has a linear factor contributed by $p$ and $i$, which is a contradiction to the fact that $i$ contributes to $f$.  So we must have at least one of $m$ and $m+1$ in $I_i$.  However, all propagators in $S$ contribute to $I_{i}$ strictly before $m$ {\color{violet} by Corrollary \ref{GN alg well defined}}, so the matrix giving $\Delta_{I_i}$ {\color{violet} where the rows appear in the order defined by $I_i$?} has the form
\[
\begin{bmatrix} A & B \\ 0 & C\end{bmatrix}
\]
where $A$ is the $|S|\times |S|$ matrix indexed by the propagators in $S$ and the vertices in $I_i(S)$. No other propagators can be supported on these vertices since all other propagators are outside of $p$, and $p$ is the first propagator supported at $i$; this explains the zero block.  Therefore $\Delta_{I_i} = \det A \det C$, and both factors are nontrivial since at least one of $m$ and $m+1$ appear in $I_i$.  Since $p\in \cP_{in}(p)$ we have $f|\det A${\color{violet}. If we remove the propagators that contribute variables to the minor $C$,}\st{ and if we remove the propagator outside of $p$ that contributes $m$ or $m+1$,} we get a diagram $W'$ with fewer propagators for which $\det A | \Delta_{I^{W'}_i}$ still holds, hence $f|\Delta_{I^{W'}_i}$.  This contradicts the minimality of our choices. \sanote{trying to make clear that the minor changes with the diagram. May have just made it too notation heavy.}
